---
title: "STAT 545A Homework 3"
author: "Tian Gao"
date: "2018/9/27"
output: github_document
---
# Bring rectangular data in 
```{r}
library(gapminder)
library(tidyverse)
```

# Get the maximum and minimum of GDP per capita for all continents.
```{r}
# get data of maximum and minimum of GDP per capita for all continents
gapminder %>% 
  # group by continent
  group_by(continent) %>% 
  # calculate summaries
  summarize(
    max_gdpPercap = max(gdpPercap),
    min_gdpPercap = min(gdpPercap),
  )
```
# Look at the spread of GDP per capita within the continents.
* First I'll show the table of min, max, mean, sd of GDP percap
```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(
    min_GDPPercap = min(gdpPercap),
    max_GDPPercap = max(gdpPercap),
    mean_GDPPercap = mean(gdpPercap),
    sd_GDPPercap = sd(gdpPercap)
  )%>%
  knitr::kable()
```


* In order to show the spread of GDP per capita within the continents, a plot should be clear and easy to understand. I put the distribution of all continents in one plot, thus the difference is obvious.
```{r}
gapminder %>% 
  # gdpPercap as x axis
  ggplot(aes(x=gdpPercap, color=continent)) +
  # scale the x axis as the variance between continents is huge
  scale_x_log10()+
  # show the distribution of data
  geom_density()
```

* Then I'll show the numbers in above table using box plot
```{r}
gapminder %>%
  ggplot(aes(x=continent, y=gdpPercap)) + 
  geom_boxplot()
```
# Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
* For this part I will calculate the weighted mean, weighting by population
```{r}
gapminder %>%
  group_by(year) %>%
  summarize(
    vanilla_mean_lifeExp = mean(lifeExp),
    weighted_mean_lifeExp = weighted.mean(lifeExp,pop)
  )%>%
  knitr::kable()
```
* It is not obvious enough shown in the above table. Thus I try to put this into a plot. 
```{r}

gapminder %>%
  group_by(year) %>%
  mutate(avgLifeExp = mean(lifeExp)) %>%
  ggplot(aes(x = continent, y = lifeExp)) + 
  geom_jitter(alpha = 0.5, aes(color = lifeExp>avgLifeExp))+ 
  facet_wrap( ~year)
```
# How is life expectancy changing over time on different continents?
* First I'll document the life expectancy in a table for each continent
* Asia
```{r}
gapminder %>%
  filter(continent == 'Asia') %>%
  group_by(year) %>%
  summarize(Avg=mean(lifeExp))
```

* Africa
```{r}
gapminder %>%
  filter(continent == 'Africa') %>%
  group_by(year) %>%
  summarize(Avg=mean(lifeExp))
```

* Americas
```{r}
gapminder %>%
  filter(continent == 'Americas') %>%
  group_by(year) %>%
  summarize(Avg=mean(lifeExp))
```

* Europe
```{r}
gapminder %>%
  filter(continent == 'Europe') %>%
  group_by(year) %>%
  summarize(Avg=mean(lifeExp))
```

* Oceania
```{r}
gapminder %>%
  filter(continent == 'Oceania') %>%
  group_by(year) %>%
  summarize(Avg=mean(lifeExp))
```

* As table is not obvious enough to show the data, I'll then draw a plot to reflect this. 
```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarize(
    mean_lifeExp = mean(lifeExp)
  )%>%
  # year as x axis and mean_lifeExp as y axis
  ggplot(aes(year, mean_lifeExp)) +
  
  # facetting by continent
  facet_wrap(~continent, scales = "free_y")+
  # make a line plot with points
  geom_line() +
  geom_point()
```

* From the plot above we can see that the life expectancy is increasing over the years.

# Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy ¨C you decide ¨C a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.
* For this question I chose mean as the benchmark

```{r}
mean(gapminder$lifeExp)
```
* Before I make a plot, I'll first show in table how to determine whether a country have a life expectancy lower that benchmark. 
```{r}
 gapminder %>% 
  group_by(country, year) %>% 
  # check if a country has low lifeExp
  mutate(
    low_lifeExp = lifeExp < mean(gapminder$lifeExp)
  ) %>% 
  select(country, year, lifeExp, low_lifeExp)%>%
  head(100)%>%
  knitr::kable()

```

* From here, we can plot based on the number of boolean value. More specificly we can count the number of appearance of TRUE. 

```{r}
 gapminder %>% 
  group_by(country, year) %>% 
  # check if a country has low lifeExp
  mutate(
    low_lifeExp = lifeExp < mean(gapminder$lifeExp)
  ) %>% 
  ggplot(aes(x=year, fill=low_lifeExp)) +
  facet_grid(~continent)+
  # make it a bar plot
  geom_bar()

 gapminder %>% 
  group_by(continent, year) %>% 
  mutate(
    low_lifeExp = lifeExp < mean(gapminder$lifeExp)
  ) %>%
  summarise(low_ct = sum(low_lifeExp))%>%
  ggplot(aes(x=year, y=low_ct)) +
  # make it a bar plot
  geom_line(aes(color = continent))+
   geom_point(aes(color = continent))

```

* The plot above is the report of absolute of countries with low life expectancy over time by continent
  
# Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.

* For this part I will try to find the relation between GDPPercap and lifeExp in .
```{r}
gapminder%>%
  filter(continent == "Asia") %>%
  # pop as x axis and lifeExp as y axis
  ggplot(aes(x=gdpPercap, y=lifeExp)) +
  # scale y axis by log10
  scale_y_log10() +
  # facetting by country
  facet_wrap(~country, scales="free") +
  # make a line plot
  geom_line() +
  # make a better x axis
  scale_x_continuous(breaks=NULL)
```