---
title: "STAT 545A Homework 3"
author: "Tian Gao"
date: "2018/9/27"
output: github_document
---
# Bring rectangular data in 
```{r}
library(gapminder)
library(tidyverse)
```

# Get the maximum and minimum of GDP per capita for all continents.
```{r}
```
# Look at the spread of GDP per capita within the continents.
```{r}
```
# Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
```{r}
```
# How is life expectancy changing over time on different continents?
```{r}
```
# Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy ¨C you decide ¨C a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.
```{r}
```
# Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.
```{r}
```